\documentclass[margin,line]{res}

\usepackage{amsmath, amssymb,graphicx}
\usepackage{hyperref}
%\oddsidemargin -.5in
%\evensidemargin -.5in
%\textwidth=6.0in
%\itemsep=0in
%\parsep=0in

\newenvironment{list1}{
  \begin{list}{\ding{113}}{%
      \setlength{\itemsep}{0in}
      \setlength{\parsep}{0in} \setlength{\parskip}{0in}
      \setlength{\topsep}{0in} \setlength{\partopsep}{0in} 
      \setlength{\leftmargin}{0.17in}}}{\end{list}}
\newenvironment{list2}{
  \begin{list}{$\bullet$}{%
      \setlength{\itemsep}{0in}
      \setlength{\parsep}{0in} \setlength{\parskip}{0in}
      \setlength{\topsep}{0in} \setlength{\partopsep}{0in} 
      \setlength{\leftmargin}{0.2in}}}{\end{list}}


\begin{document}

\name{Laura Florescu \vspace*{.1in}}

\begin{resume}
\section{\sc Contact Information}
\vspace{.05in}
\begin{tabular}{@{}p{2in}p{4in}}
801 Corbett Ave      & \url{http://cims.nyu.edu/~florescu} \\            
San Francisco, CA 94131    & {\it E-mail:}  lara.florescu@gmail.com\\
{\it Phone:}  (503) 470 0184 & {\it GitHub:}  \url{https://github.com/florescl} \\       
\end{tabular}
{\it Linkedin:} \url{https://www.linkedin.com/in/laura-florescu-48692b103}  


\section{\sc Experience}
{\bf MosaicML}, San Francisco, CA \hfill{2021 - present }\\ 
AI Researcher. Projects include \\
- algorithmic speedups of neural network training \\
- segmentation, detection model implementation \\
- investigation of alternative custom hardware

{\bf Sambanova Systems}, Palo Alto, CA \hfill{2019 - 2021 }\\
\vspace{0.cm}
Principal Engineer in ML team. Responsibilities including \\
- bring-up of various models in NLP, GAN, CNN \\
- helping design future chips \\
- benchmarking models 
\vspace{-0.1cm}

{\bf Groq}, Menlo Park, CA \hfill{2017 - 2019 }\\
\vspace{0.cm}
Software engineer in Compiler Team working on AI hardware accelerators. Projects include \\
- efficiently compiling and benchmarking classical and custom DL models \\
- graph rewrites and optimizations \\
- IR primitives lowering \\
- leading compiler testing
\vspace{-0.1cm}

{\bf Microsoft Research}, Redmond, WA \hfill {2013 - 2016}\\
\vspace{-.1cm}
Algorithms and stochastic models R\&D

{\bf Los Alamos National Laboratory}, Los Alamos, NM  \hfill 2011 - 2012\\
\vspace{-.1cm}
Computational plasma physics R\&D

{\bf Oracle}, Portland, OR \hfill 2008 - 2010\\
\vspace{-.1cm}
Oracle Database Design and Programming SQL design
\vspace{-.2cm}


\section{\sc Education}
{\bf New York University, Courant Inst. of Mathematical Sciences}, New York
\begin{list2}
\item Ph.D., Computer Science, 2012-2017. Advisor: Joel Spencer
\item Thesis: Random growth models, \textbf{Dean Dissertation Fellowship award}.

\vspace*{-.1in}
\end{list2}

{\bf Reed College}, Portland, OR USA.
\begin{list2}
\item B.A., Mathematics  May, 2011, {\it Phi Beta Kappa}.
\vspace*{-.05in}
\end{list2} 

\section{\sc Honors and Awards}
UnifyID AI Fellowship, 2016-2017.\\
Dean's Dissertation Fellowship, NYU, 2016-2017.\\
Summer Research Fellowship, NYU CS Department, 2013, 2014, 2016.\\
Henry MacCracken Fellowship, NYU, Courant Institute. 2012-2017.\\
Phi Beta Kappa, 2011.\\
Reed College Commendation for Excellence in Scholarship, 2007 - 2011. \\
\emph{1st} prize, Database Design International Contest, Oracle Academy, 2007.

\section{\sc Academic and Industry Visits}
{Simons Institute for the Theory of Computing}, Berkeley, CA. Spring 2016. \\
{Microsoft Research, Theory Group}, Redmond, WA. short visits, '12-'15, Spring 15. \\
{Institute for Mathematics and Applications}, Minneapolis, MN, Fall 2014.
\vspace*{-.2in}
\\

\section{\sc Book}
Spencer, J. with Florescu, L. \emph{Asymptopia}, \url{http://www.ams.org/bookstore-getitem/item=stml-71}, Student Mathematical Library, New York, NY, 2014.

\section{\sc Teaching experience}
Fundamental algorithms, NYU-Courant \hfill {Fall 2015, Spring 2016}\\
Asymptopia, NYU-Courant \hfill {Spring 2014}

\section{\sc Papers}
Florescu, L., et al. \emph{Accelerating Scientific Applications With SambaNova Reconfigurable Dataflow Architecture}, IEEE 2021. \\
Florescu, L., et al. \emph{Optimal control for diffusions on graphs}, SIDMA, 2018. \\
Florescu, L., et al. \emph{Spectral thresholds in the bipartite SBM}, COLT16.\\
Florescu, L., et al. \emph{The range of rotor walk}, Amer. Math Monthly, 2016\\
Florescu, L., et al. \emph{Escape rates for rotor walk in $\mathbb{Z}^d$}, SIDMA, 2014.\\ 
Florescu, L., et al. \emph{Heat diffusion with frozen boundary}, J. Stat Phys, 2015. \\
Florescu, L., et al. \emph{Dominos and Sandpiles}, EJC, 22(1), 2015.

\section{\sc Selected Talks}

\emph{Making neural networks train faster algorithmically}, GGX, 2022 \\
\emph{Spectral Thresholds in the Bipartite Stochastic Block Model}, NYU, Paris, Berkeley, 2016\\
\emph{Optimal control for diffusions on a graph}, Casa Oaxaca, Mexico, 2015\\
\emph{The range of a deterministic random walk}, Reed College, Math Colloquium, 2015\\
\emph{The range of rotor walk}, Rutgers, Georgia Tech, Eindhoven 20115\\
\emph{Frozen random walk}, IMA, University of Minnesota, 2014\\

\vspace*{-5.5mm}
\section{\sc Professional Memberships and Service}
Organizer for ValleyML AI Hardware conference \\
Girls who code Facilitator \\
Reviewer for Women in Machine Learning Workshop\\
Referee for SODA, SIAM J of Discrete Math, Electronic Comm. in Probability, Journal of Inequalities and Applications \\
Mathematical Reviewer, American Mathematical Society \\
Oracle International Database Design Competition Judge, 2009 and 2010\\
\vspace*{-5.5mm}

\section{\sc Skills} 
Programming:  Pytorch, Tensorflow, Haskell, Python, C, Unix, Mathematica. \\
Languages: Fluent in Romanian; intermediate French, German.
\vspace{-.2cm}


\end{resume}
\end{document}




